---
title: ""
author: "Tony Carilli"
date: '`r format(lubridate::today(), "%B %d, %Y")`'
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA,
                      options(digits = 3, scipen = 999),
                      eval = FALSE,
                      message = FALSE)
library(tidyverse)
```

# APIs and Descriptions 

## U.S. State and Territorial Gathering Bans: March 11-December 31, 2020 by County by Day

<https://data.cdc.gov/Policy-Surveillance/U-S-State-and-Territorial-Gathering-Bans-March-11-/7xvh-y5vh>

```{r}
## U.S. State and Territorial Gathering Bans: March 11-December 31, 2020 by County by Day
gathering_bans <- 
  jsonlite::fromJSON("https://data.cdc.gov/resource/7xvh-y5vh.json")
```


## U.S. State and Territorial Stay-At-Home Orders: March 15 -- December 31 by County by Day

<https://data.cdc.gov/Policy-Surveillance/U-S-State-and-Territorial-Stay-At-Home-Orders-Marc/y2iy-8irm>

```{r}
## U.S. State and Territorial Stay-At-Home Orders: March 15 – December 31 by County by Day
jsonlite::fromJSON("https://data.cdc.gov/resource/y2iy-8irm.json")
```

## U.S. State and Territorial Public Mask Mandates From April 10, 2020 through April 28, 2021 by County by Day

<https://data.cdc.gov/Policy-Surveillance/U-S-State-and-Territorial-Public-Mask-Mandates-Fro/62d6-pm5i>

```{r}
jsonlite::fromJSON("https://data.cdc.gov/resource/62d6-pm5i.json")
```

## U.S. State and Territorial Orders Closing and Reopening Bars Issued from March 11, 2020 through December 31, 2020 by County by Day

<https://data.cdc.gov/Policy-Surveillance/U-S-State-and-Territorial-Orders-Closing-and-Reope/9kjw-3miq>

```{r}
jsonlite::fromJSON("https://data.cdc.gov/resource/9kjw-3miq.json")
```

## U.S. State and Territorial Orders Closing and Reopening Restaurants Issued from March 11, 2020 through December 31, 2020 by County by Day

<https://data.cdc.gov/Policy-Surveillance/U-S-State-and-Territorial-Orders-Closing-and-Reope/azmd-939x>

```{r}
jsonlite::fromJSON("https://data.cdc.gov/resource/azmd-939x.json") %>% 
  select(-1:-5) %>% 
  select(-contains("date")) %>% 
  unique()
```


# Read CSV Files

## Gathering bans

```{r gathering-code}
# gathering_bans <-
  here::here(
    "data",
    "U.S._State_and_Territorial_Gathering_Bans__March_11-December_31__2020_by_County_by_Day.csv"
  ) %>% 
  read_csv() %>% 
  janitor::clean_names() %>% 
  distinct(general_gb_order_group, general_gb_order_code) %>% 
  arrange(general_gb_order_code)
```

```{r}
gathering_bans %>% 
  distinct(general_gb_order_group, general_gb_order_code) %>% 
  arrange(general_gb_order_code)
```


```{r gathering-bans, eval=TRUE}
gathering_bans <- 
 here::here(
    "data",
    "U.S._State_and_Territorial_Gathering_Bans__March_11-December_31__2020_by_County_by_Day.csv"
  ) %>% 
  read_csv(show_col_types = FALSE) %>% 
  janitor::clean_names() %>% 
  mutate(fips = str_c(fips_state, fips_county)) %>% 
  select(
    state = state_tribe_territory,
    county = county_name,
    fips,
    date,
    general_gb_order_code
    ) %>% 
  mutate(gathering_ban = ifelse(general_gb_order_code == 1, 0 , 1)) %>% 
  # group_by(fips) %>% 
  # mutate(days = mean(gathering_ban)) %>% 
  # summarize(mean(days))
  # filter(state == "VA") %>% 
  group_by(fips) %>% 
  mutate(gathering_days = sum(gathering_ban)) %>% 
  ungroup() %>% 
  select(state, county, fips, gathering_days) %>% 
  distinct(fips, .keep_all = TRUE)
```

## Stay at Home orders

```{r stay-at-home-data}
here::here(
  "data", 
  "U.S._State_and_Territorial_Stay-At-Home_Orders__March_15___December_31_by_County_by_Day.csv"
  ) %>% 
  read_csv() %>% 
  janitor::clean_names() %>% 
  distinct(current_order_status, order_code)

# %>%
#   mutate(fips = str_c(fips_state, fips_county)) %>% 
#   select(state = state_tribe, count_name, fips,
#          current_order_stats, order_code)
```
Code any mandatory as 1 No order and advisory as 0

```{r stay-at-home-orders, eval=TRUE}
stay_at_home_orders <- 
here::here(
  "data", 
  "U.S._State_and_Territorial_Stay-At-Home_Orders__March_15___December_31_by_County_by_Day.csv"
  ) %>% 
  read_csv(show_col_types = FALSE) %>% 
  janitor::clean_names() %>% 
   mutate(fips = str_c(fips_state, fips_county)) %>% 
  mutate(stay_at_home_days = 
           case_when(
             order_code == 7 ~ 0,
             order_code == 1 ~ 1,
             order_code == 6 ~ 0,
             order_code == 2 ~ 1,
             order_code == 3 ~ 1,
             order_code == 5 ~ 1
           )) %>% 
  group_by(fips) %>% 
  mutate(stay_at_home_days = sum(stay_at_home_days)) %>% 
  ungroup() %>% 
  select(state = state_tribe_territory, county = county_name, fips, stay_at_home_days) %>% 
  distinct(fips, .keep_all = TRUE)
```



## Masks

```{r}
mask_days <- 
here::here(
  "data",
  "U.S._State_and_Territorial_Public_Mask_Mandates_From_April_10__2020_through_April_28__2021_by_County_by_Day.csv"
) %>%
  read_csv() %>%
  janitor::clean_names() %>% 
  mutate(date = as.Date(lubridate::mdy_hms(date))) %>% 
  distinct(order_code, current_order_status)
```


```{r mask-mandates, eval=TRUE}
mask_mandates <- 
here::here(
  "data",
  "U.S._State_and_Territorial_Public_Mask_Mandates_From_April_10__2020_through_April_28__2021_by_County_by_Day.csv"
) %>%
  read_csv(show_col_types = FALSE) %>%
  janitor::clean_names() %>%
  mutate(date = as.Date(lubridate::mdy_hms(date)),
    fips = str_c(
      str_pad(as.character(fips_state), 2, "left", "0"),
      str_pad(as.character(fips_county), 3, "left", "0")
    ),
    order_code = if_else(order_code == 2, 0, 1)
    ) %>% 
  filter(date < "2021-01-01") %>% 
  group_by(fips) %>% 
  mutate(mask_days = sum(order_code)) %>% 
  ungroup() %>% 
  select(state=state_tribe_territory, county = county_name, fips, mask_days) %>% 
  distinct(fips, .keep_all = TRUE)
```


## Bars

```{r bar-closures, eval=TRUE}
bar_closures <- 
here::here(
  "data",
  "U.S._State_and_Territorial_Orders_Closing_and_Reopening_Bars_Issued_from_March_11__2020_through_December_31__2020_by_County_by_Day.csv"
) %>% 
  read_csv(show_col_types = FALSE) %>% 
  janitor::clean_names() %>% 
  mutate(
    date = as.Date(lubridate::mdy_hms(date)),
    fips = str_c(
      str_pad(as.character(fips_state), 2, "left", "0"),
      str_pad(as.character(fips_county), 3, "left", "0")
    ),
    order_code = if_else(order_code %in% c(1, 2, 3), 0, 1)
  ) %>% 
  group_by(fips) %>% 
  mutate(bar_days = sum(order_code)) %>% 
  ungroup() %>% 
  select(state = state_tribe_territory, county = county_name, fips, bar_days) %>% 
  distinct(fips, .keep_all = TRUE)
  
```

## Restaurants

```{r restaurant-closures, eval=TRUE}
restaurant_closures <- 
here::here(
  "data",
  "U.S._State_and_Territorial_Orders_Closing_and_Reopening_Restaurants_Issued_from_March_11__2020_through_December_31__2020_by_County_by_Day.csv"
) %>%
  read_csv(show_col_types = FALSE) %>%
  janitor::clean_names() %>%
  mutate(
    date = as.Date(lubridate::mdy_hms(date)),
    fips = str_c(
      str_pad(as.character(fips_state), 2, "left", "0"),
      str_pad(as.character(fips_county), 3, "left", "0")
    ),
    order_code = if_else(order_code %in% c(1, 2, 3), 0, 1)
  ) %>%
  group_by(fips) %>% 
  mutate(restaurant_days = sum(order_code)) %>% 
  ungroup() %>% 
  select(state = state_tribe_territory, county = county_name, fips, restaurant_days) %>% 
  distinct(fips, .keep_all = TRUE)
```

# Create restrictions data

```{r restriction-data, eval=TRUE}
bar_closures %>% 
  left_join(gathering_bans) %>% 
  left_join(mask_mandates) %>% 
  left_join(restaurant_closures) %>% 
  left_join(stay_at_home_orders) %>% 
  write_csv(file = "data/restriction_days.csv")
```

